---
title: 'Exercise 4: Base R vs. Tidyverse'
author: "Natalie Jones"
date: "10/15/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
getwd()
```


# Base R tasks

1. Download the food_coded.csv file 
2. Load the CSV file into your R environment.

```{r}
food <- read.csv("food_coded.csv")
```


3. Extract the first 95 rows.

```{r}
foodextraction <- food[1:95,]
head(foodextraction)
```

4. Look at the following variables using both name and column index/number.

    * GPA
    * calories_chicken
    * drink
    * fav_cuisine
    * father_profession
    * mother_profession
    
```{r}
a<- food[food$GPA==3.3 ,]
b<-as.data.frame(food[, 'GPA'])

food[, 'GPA']
food[, 1]

food[, 'calories_chicken']
food[, 4]


food[, 'drink']
food[,16 ]

indexoffood <- foodextraction[,c(1,4,16,26,25,45)]


```

5. Create a new variable for how healthy each person feels but convert the scale from 1 to 10 to 1 to 100.

```{r}
food$healthy_feeling
food.new <- food
food.new$healthy_feeling <- (food$healthy_feeling)*10
head(food.new$healthy_feeling)
```

6. Filter to students who are female and have GPAs that are above 3.0.

```{r}
gpa_female <- food[food$Gender=="1" & food$GPA > 3.0,]
head(gpa_female)

```

7. Find the mean and standard deviation for the following variables, and summarize them in a data frame.

    * chicken_calories
    * tortilla_calories
    * turkey_calories
    * waffle_calories
    
    
```{r}
food.new <- food[,c(4,54,55,60)]
m <- sapply(food.new, mean, na.rm = T)
sd <- sapply(food.new, sd, na.rm = T)
calories_m_md <- rbind (m, sd)
head(calories_m_md)
```
8. Summarize GPA and weight within the gender and cuisine variables.

```{r}
class(food$weight)
food$weight <- as.numeric(food$weight)
food$GPA <- as.numeric(food$GPA)

food_men <- food[food$Gender ==2,]
food_women <- food[food$Gender ==1,]

gpa_mean_men <- tapply(food_men$GPA, food_men$cuisine, mean, na.rm = T)
gpa_sd_men <- tapply(food_men$GPA, food_men$cuisine, sd, na.rm =T)

gpa_mean_women <- tapply(food_women$GPA, food_women$cuisine, mean, na.rm = T)
gpa_sd_women <- tapply(food_women$GPA, food_women$cuisine, sd, na.rm = T)

weight_mean_men <- tapply(food_men$weight, food_men$cuisine, mean, na.rm = T)
weight_sd_men <- tapply(food_men$weight, food_men$cuisine, mean, na.rm = T)

weight_mean_women <- tapply(food_women$weight, food_women$cuisine, mean, na.rm = T)
weight_sd_women <- tapply(food_women$weight, food_women$cuisine, mean, na.rm = T)
```

# Tidyverse tasks

1. Download the facebook-fact-check.csv

```{r}
fb_fact <- read.csv ("facebook-fact-check.csv")
library(tidyverse)


```

2. Load the CSV file into your R environment.

3. Extract the last 500 rows.

```{r}
fb_extract <- fb_fact %>% slice_tail(n = 500)
head(fb_extract)
```


    Hint: Check out the [top_n() page](https://rdrr.io/github/YTLogos/dplyr/man/top_n.html) to figure out how to extract the last 500 rows instead of the first 500 rows.
    
4. Look at the even-numbered column indices only. Identify them by name.

```{r}
row_odd <- seq_len(nrow(fb_extract)) %% 2
data_row_odd <- fb_extract[row_odd == 1, ]
colnames(data_row_odd)
```

5. Using `mutate`, create a new variable called `post_type_coded` that renames each post type to the following:

    * link = 1
    * photo = 2
    * text = 3
    * video = 4
    
    Hint: look up case_when within tidyverse. You can also use if_else
    
    
```{r}
fb_extract <- fb_extract %>%
    mutate(post_type_coded = ifelse("Post.Type" == "link", '1',
                                    ifelse("Post.Type" == 'photo', '2',
                                           ifelse("Post.Type" == 'text', '3', '4'))))
```
    
6. Arrange page names in reverse order.

```{r}
fb_extract <- fb_extract %>% arrange(desc(Page))
head(fb_extract)

```
    
7. Find the mean and standard deviation for the following variables, and summarize them.

    * share_count
    * reaction_count
    * comment_count

```{r}
fb_fact %>%
    summarise(share_count = mean(share_count, na.rm=T),
        reaction_count = mean(reaction_count, na.rm=T),
        comment_count = mean(comment_count, na.rm=T))
        
```
8. Summarize the mean and standard deviations in Question 7 with the "mainstream" values in the `category` variable.

```{r}
fb_main <- fb_fact %>%
    filter(Category == "mainstream") %>%
    summarize(share_count_m = mean(share_count, na.rm=T), 
        reaction_count_m = mean(reaction_count, na.rm =T),
        comment_count_m = mean(comment_count, na.rm=T),
        share_count_sd = sd(reaction_count, na.rm=T), 
        reaction_count_sd = sd(reaction_count, na.rm=T),
        comment_count_sd = sd(comment_count, na.rm=T))
```

# Submit

Email me (laaker@wisc.edu) the link to your `ps811-exercises` repository when you are done.